\chapter{Introduction}

This paper describes the general design of the gui2 project. The paper is
divided in several chapters:
\begin{description}
\item[Chapter 1] Explains why yet another document is added instead of using the
	doxygen and a short history of the gui2 project.
\item[Chapter 2] Goes into the general design and structure of the gui2 project.
\item[Chapter 3] Dives deeper into certain parts of the design of certain
	classes and algorithms.
\end{description}

A omission in this paper is the simple story of how to add your own widgets in
the C++ code\footnote{The WML part is described in the wiki.}. I have a slightly
outdated paper on the subject, but decided not to add it; the paper needs
polishing and updates, but I've some design changes planned for the 1.9
development series. These changes will also change the way of adding new
widgets, so that paper needs to be rewritten and added after the rewrite is
complete.

\section{Why why why?}

Why is this document written in \LaTeX{} instead of doxygen? The reason is
simple; during the development of the iterator class I felt I had to write a
lot. Writing much documentation in doxygen feels awkward. A lot of time is spend
on formatting, the number of subsections is limited (too low in my opinion).
Also it's not possible to document why certain functions aren't implemented, a
solution would be to make the function private and declared only, but that feels
like an ugly hack.

Why is the document needed? In my opinion most manuals written in doxygen are
great as reference manuals but fail to explain how to do certain things with the
library and fail to explain why certain things are implemented in the way it is.
Writing a separate document separated from the code has the disadvantage of
getting earlier out of sync when an implementation details differs, but this can
be used as an advantage as well. Since it's hard to keep the details in sync,
it's a good reason not to dive into these details. This makes a natural
separation between which part should be in doxygen and which part in this
manual.

Why after the above, are some parts of gui2 documented in separate pages in
doxygen? This indeed seems a contradiction, but it's not. Remember the first
paragraph stating it's annoying to write this kind of documentation in doxygen?
These parts are where I realized it's annoying to do it in doxygen and haven't
been incorporated in this document\footnote{These parts are still mentioned in
this paper, but they refer to the doxygen documentation.}.

\section{History}

The gui2 project was started late in 2007 to fix several problems I had with the
current gui at that time. The problems I had were:
\begin{enumerate}
\item When using small resolutions the gui started to look bad, since widgets
	simply left their container and were drawn outside it.
\item Most of the gui was hard-coded in the C++ code and thus not configurable
	for normal users.
\item ThemeWML, the part that allows the user to configure the layout, is not
	well understood by the current developers.
\end{enumerate}

This let to the following design goals:

\begin{enumerate}
\item The gui should look well at every possible resolution and automatically
	fit well.
\item Everything, or at least as much as possible should be configurable by WML.
\item The project needed more documentation as the current ThemeWML, and explain
	how to create your own guis.
\end{enumerate}

At the time of this writing the project is still work in progress\footnote{It
was never expected to be a short during project.} and will still take a while to
finish. The initial goals were pretty clear, and still are. On the other hand
how to implement certain parts turned out to be less clear.

The library is the first time I designed a gui toolkit and like with all larger
projects, you need to learn from mistakes; preferably of others, but you're doomed
to make your own as well. The design process is an iterative one, causing some
parts of the design to be changed several times. At the time of
writing\footnote{Shortly before the 1.8 release.} a lot of areas feel stable,
but others are still planned to be rewritten. Mainly the handling of scrollbar
widgets feels awkward and hacky and are slated for a redesign.

Other parts of the gui are still in the planning stage and not implemented yet,
but the project moves along at a steady pace and more features will be added in
the future.

\paragraph{}

The history of this document also goes back a long time, it's mostly build from
various scraps of notes I have on my system. The plan was to change these notes
to doxygen format some day put it in the source. The problem was to find a nice
place in the source to do so.

This lead to the fact the notes just stayed on my system and the number of
notes accumulated. During travelling to work I started to design some things for
the 1.9 development series and for the iterator class I needed a new kind of
document. I decided that was the final straw and started to work on writing this
document while travelling.

Since you read this document, it's uploaded in the Wesnoth source tree. However, this
doesn't mean the document is complete. In fact it means more of the opposite, the
draft is uploaded and I can start to work on finishing the document by adding
the missing information.
